---
title: "BaileyLabMRRs"
author: "James Wade"
date: "November 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The programs within this repository are designed to process and analyze data 
from the M1 Maverick Detection System (Genalyte, Inc., San Diego, CA). Details 
about each program are provided within each program file. The instructions are 
also reproduced below. Sample data is available in the SampleData folder within 
this repository. 

***

## Getting started

###To get started, follow these steps:

1. If you have not already done so, download and install R and RStudio. There 
are many online tutorials available with instructions.
2. Ensure all of your packages are up to date.
3. Download this respository from Github. From [this link](https://github.com/BaileyLabUM/BaileyLabMRRs), click on
`Clone or Download` and then click `Download Zip`. Alternatively [click here](https://github.com/BaileyLabUM/BaileyLabMRRs/archive/master.zip) to 
download the repository directly.
4. Unzip the repository into a directory your choosing.
5. Open RStudio, and set the working directory to the directory with the
program files and sample data. See instructions on setting the working directory
in R [here](https://www.statmethods.net/interface/workspace.html). For example,
if you are using a Windows machine and downloaded into your `Downloads` folder,
you could set your working directory by executing the following line in the
console: `setwd("C:/Users/USERNAME/Download/BaileyLabMRRs-master")`. Depending
on how you extracted the compressed files you may need to use `setwd("C:/Users/USERNAME/Download/BaileyLabMRRs-master/BaileyLabMRRs-master")`
instead.

***

##Library Specifics

The purpose of this program is to process with the raw data from the
Maverick M1 detection system (Genalyte, Inc., San Diego, CA) and output simple
line graphs, bar charts, and box plots. The functions also generate companion
csv files containning processed the prcoessed data for subsequent analysis.
The folder containing output from the M1 typically consists of:

1. a csv file for each ring and
2. a comments file the describes the experimental run

The comments file is not needed for this program. In addition to the csv files
for each ring, a separate file containing the chip layout is required. An
example of a chip layout file is provided in the "BaileyLabMRRs" repository
located at https://github.com/BaileyLabUM/BaileyLabMRRs. See
"groupNames_allClusters.csv" for an example.

Note: This version of the software is optimized for the Bailey lab's HRP
assay. See dx.doi.org/10.1021/acscentsci.5b00250 for a description. However,
input variables can be altered to accomodate many alternative experiments.

The following libraries are used in this program: tidyverse, ggthemes.
To install these packages, run the following code in the console:
      install.packages(c("tidyverse", "ggthemes"))

To use the program:

1. Ensure the you have the necessary libraries installed. See note above for instructions on installing libraries.
2. Copy the chip layout file (e.g., "groupNames_XPP.csv") into the directory containing the raw ring data.
     Note: This program has the highest chance of success if the directoy _only_ contains:  
     1. raw ring data files (e.g., "03.csv", "04.csv", etc.) and
     2. the chip layout file (e.g., "groupNames_XPP.csv")
3. Source all of the code from this file. There are multiple ways, but one method is to click the `Source` button on the window of the Source Console in RStudio after opening this file in RStudio.

4. Set the working directory to the folder containing the raw data and chip layout for example ```setwd("C:/Users/USERNAME/Documents/CHIPNAME_gaskGASK_DATE")```

5. Execute the code by running the AnalyzeData function. This function requires 5 input variable:
    1 filename - the name of the chip layout file
    2 loc - the name of folder to store generated files
    3 fsr - a logical variable the corrects for free spectral range shifts; most likely leave as FALSE if you don't know what this is
    4 time1 - the later time for net shift measurements
    5 time2 - the earlier time for net shift measurments
Note: to calculate net shift measurements, the relative shift at time2 is subtracted from time1 (netshift = time1 - time2).  
Here is an example of code to run: 
```{r, eval = F}
analyzeData(filename = "groupNames_XPP.csv",
            time1 = 51,
            time2 = 39,
            loc = "plots",
            fsr = FALSE, 
            chkRings = FALSE)
```
